### 2020-11-05 submission/manuscript.tex
Figure legends and figure edits

\documentclass[]{article}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
% Redefine \includegraphics so that, unless explicit options are
% given, the image width will not exceed the width of the page.
% Images get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\ScaleIfNeeded{%
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother
\let\Oldincludegraphics\includegraphics
{%
 \catcode`\@=11\relax%
 \gdef\includegraphics{\@ifnextchar[{\Oldincludegraphics}{\Oldincludegraphics[width=\ScaleIfNeeded]}}%
}%
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}

\author{}
\date{\vspace{-2.5em}}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\modulolinenumbers
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}

@@

\section{Removal of rare sequences from 16S rRNA gene sequence surveys
biases the interpretation of community structure
data}\label{removal-of-rare-sequences-from-16s-rrna-gene-sequence-surveys-biases-the-interpretation-of-community-structure-data}

@@

Patrick D. Schloss${^\dagger}$

@@

$\dagger$ To whom correspondence should be addressed:

@@

\subsection{Research article format}\label{research-article-format}

@@

\subsection{Abstract}\label{abstract}

@@

\subsection{Importance}\label{importance}

@@

\subsection{Results}\label{results}

@@

\textbf{The nature of singletons.} Removal of rare ASVs is commonly
justified as a method of removing ASVs that are artifacts. If such ASVs
are artifacts, then one would expect the number of singleton ASVs to
accumulate with sequencing depth. Contrary to this expectation, the
median percentage of sequences that were discarded when singleton ASVs
were removed from each dataset varied between 0.42 and 22.23\%
(bioethanol and seagrass). In addition, with the exception of the
samples from the marine and sediment datasets (Spearman correlation,
P\textgreater{}0.05), the fraction of singleton ASVs in samples was
negatively correlated with the number of sequences in each sample with a
range between -0.27 and -0.87 (rice and bioethanol) (Figure 1A). This
showed that with additional sequencing, the probability of seeing
singleton ASVs in multiple samples was greater than the probability of
generating an artifact. This suggests that the singleton ASVs are not as
likely to be artifacts as previously thought. Furthermore, if singleton
ASVs were artifacts, then one would not expect to find them in other
samples from the same dataset. In fact, singleton ASVs from samples with
fewer sequences were often found in samples with more sequences. At
least 50\% of the singleton ASVs found in the samples from the mice,
rice, seagrass, and stream datasets were found in another sample from
the same dataset (Figure 1B). Considering the likelihood of finding an
ASV duplicated in another sample is confounded by the number of samples
and inter-sample diversity, the high coverage of singleton ASVs in these
datasets was remarkable. The correlation between the number of sequences
in a sample and the fraction of that sample's singleton ASVs that were
covered by another sample in the dataset was significant and negative
for 9 of the datasets ranging between -0.31 and -0.84 for the rice and
seagrass datasets, respectively (Figure 1C). The negative correlation
indicated that the singleton ASVs in the smaller samples were more
likely to be covered by ASVs in the larger samples. Among the three
datasets without a significant correlation (Spearman correlation,
P\textgreater{}0.05), the marine and soil datasets had the fewest
samples in our collection and the stream dataset already had a high
level of coverage regardless of the number of sequences. Contrary to the
common motivation for removing rare ASVs, these results indicate that
this practice disproportionately impacts samples with fewer sequences
and likely removes more non-artifact ASVs than those that are artifacts.

@@

\subsection{Discussion}\label{discussion}

@@

\subsection{Materials and Methods}\label{materials-and-methods}

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  sequencing pipeline description
\end{itemize}

\subsection{Acknowledgements}\label{acknowledgements}

@@

\subsection{References}\label{references}

@@

\textbf{Figure 1.}

@@

\textbf{Figure 2.}

1. \textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter NT},
\textbf{Highlander SK}, \textbf{Schloss PD}. 2013. Development of a
dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq Illumina sequencing platform.
Applied and environmental microbiology \textbf{79}:5112--5120.

2. \textbf{Schloss PD}, \textbf{Westcott SL}, \textbf{Ryabin T},
\textbf{Hall JR}, \textbf{Hartmann M}, \textbf{Hollister EB},
\textbf{Lesniewski RA}, \textbf{Oakley BB}, \textbf{Parks DH},
\textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres B},
\textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{http://dx.doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.

3. \textbf{Morris A}, \textbf{Beck JM}, \textbf{Schloss PD},
\textbf{Campbell TB}, \textbf{Crothers K}, \textbf{Curtis JL},
\textbf{Flores SC}, \textbf{Fontenot AP}, \textbf{Ghedin E},
\textbf{Huang L}, \textbf{Jablonski K}, \textbf{Kleerup E},
\textbf{Lynch SV}, \textbf{Sodergren E}, \textbf{Twigg H}, \textbf{Young
VB}, \textbf{Bassis CM}, \textbf{Venkataraman A}, \textbf{Schmidt TM},
\textbf{Weinstock GM}. 2013. Comparison of the respiratory microbiome in
healthy nonsmokers and smokers. American Journal of Respiratory and
Critical Care Medicine \textbf{187}:1067--1075.
doi:\href{http://dx.doi.org/10.1164/rccm.201210-1913oc}{10.1164/rccm.201210-1913oc}.

4. \textbf{Bokulich NA}, \textbf{Subramanian S}, \textbf{Faith JJ},
\textbf{Gevers D}, \textbf{Gordon JI}, \textbf{Knight R}, \textbf{Mills
DA}, \textbf{Caporaso JG}. 2013. Quality-filtering vastly improves
diversity estimates from illumina amplicon sequencing. Nature Methods
\textbf{10}:57--59.
doi:\href{http://dx.doi.org/10.1038/nmeth.2276}{10.1038/nmeth.2276}.

5. \textbf{Fadrosh DW}, \textbf{Ma B}, \textbf{Gajer P},
\textbf{Sengamalay N}, \textbf{Ott S}, \textbf{Brotman RM},
\textbf{Ravel J}. 2014. An improved dual-indexing approach for
multiplexed 16S rRNA gene sequencing on the illumina MiSeq platform.
Microbiome \textbf{2}:6.
doi:\href{http://dx.doi.org/10.1186/2049-2618-2-6}{10.1186/2049-2618-2-6}.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
  \usepackage{amssymb}
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1.0in]{geometry}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{setspace}
\doublespacing
\usepackage[left]{lineno}
\linenumbers
\modulolinenumbers
\usepackage{helvet} % Helvetica font
\renewcommand*\familydefault{\sfdefault} % Use the sans serif version of the font
\usepackage[T1]{fontenc}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\author{}
\date{\vspace{-2.5em}}

@@

\hypertarget{removal-of-rare-sequences-from-16s-rrna-gene-sequence-surveys-biases-the-interpretation-of-community-structure-data}{%
\section{Removal of rare sequences from 16S rRNA gene sequence surveys
biases the interpretation of community structure
data}\label{removal-of-rare-sequences-from-16s-rrna-gene-sequence-surveys-biases-the-interpretation-of-community-structure-data}}

@@

Patrick D. Schloss\({^\dagger}\)

@@

\(\dagger\) To whom correspondence should be addressed:

@@

\hypertarget{research-article-format}{%
\subsection{Research article format}\label{research-article-format}}

@@

\hypertarget{abstract}{%
\subsection{Abstract}\label{abstract}}

@@

\hypertarget{importance}{%
\subsection{Importance}\label{importance}}

@@

\hypertarget{results}{%
\subsection{Results}\label{results}}

@@

\textbf{The nature of singletons.} Removal of rare ASVs is commonly
justified as a method of removing ASVs that are artifacts. If such ASVs
are artifacts, then one would expect the number of singleton ASVs to
accumulate with sequencing depth. Contrary to this expectation, the
median percentage of sequences that were discarded when singleton ASVs
were removed from each dataset varied between 0.42 and 22.23\%
(bioethanol and seagrass). In addition, with the exception of the
samples from the marine and sediment datasets (Spearman correlation,
P\textgreater0.05), the fraction of singleton ASVs in samples was
negatively correlated with the number of sequences in each sample with a
range between -0.27 and -0.87 (rice and bioethanol) (Figure 1A). This
showed that with additional sequencing, the probability of seeing
singleton ASVs in multiple samples was greater than the probability of
generating an artifact. This suggests that the singleton ASVs are not as
likely to be artifacts as previously thought. Furthermore, if singleton
ASVs were artifacts, then one would not expect to find them in other
samples from the same dataset. In fact, singleton ASVs from samples with
fewer sequences were often found in samples with more sequences. At
least 50\% of the singleton ASVs found in the samples from the mice,
rice, seagrass, and stream datasets were found in another sample from
the same dataset (Figure 1B). Considering the likelihood of finding an
ASV duplicated in another sample is confounded by the number of samples
and inter-sample diversity, the high coverage of singleton ASVs in these
datasets was remarkable. The correlation between the number of sequences
in a sample and the fraction of that sample's singleton ASVs that were
covered by another sample in the dataset was significant and negative
for 9 of the datasets ranging between -0.31 and -0.84 for the rice and
seagrass datasets, respectively (Figure 1C). The negative correlation
indicated that the singleton ASVs in the smaller samples were more
likely to be covered by ASVs in the larger samples. Among the three
datasets without a significant correlation (Spearman correlation,
P\textgreater0.05), the marine and soil datasets had the fewest samples
in our collection and the stream dataset already had a high level of
coverage regardless of the number of sequences. Contrary to the common
motivation for removing rare ASVs, these results indicate that this
practice disproportionately impacts samples with fewer sequences and
likely removes more non-artifact ASVs than those that are artifacts.

@@

\hypertarget{discussion}{%
\subsection{Discussion}\label{discussion}}

@@

\hypertarget{materials-and-methods}{%
\subsection{Materials and Methods}\label{materials-and-methods}}

\begin{itemize}
\tightlist
\item
  sequencing pipeline description
\end{itemize}

\hypertarget{acknowledgements}{%
\subsection{Acknowledgements}\label{acknowledgements}}

@@

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\leavevmode\hypertarget{ref-Kozich2013}{}%
\CSLLeftMargin{1. }
\CSLRightInline{\textbf{Kozich JJ}, \textbf{Westcott SL}, \textbf{Baxter
NT}, \textbf{Highlander SK}, \textbf{Schloss PD}. 2013. {Development of
a dual-index sequencing strategy and curation pipeline for analyzing
amplicon sequence data on the MiSeq Illumina sequencing platform}.
Applied and environmental microbiology \textbf{79}:5112--5120.}

\leavevmode\hypertarget{ref-Schloss2009}{}%
\CSLLeftMargin{2. }
\CSLRightInline{\textbf{Schloss PD}, \textbf{Westcott SL},
\textbf{Ryabin T}, \textbf{Hall JR}, \textbf{Hartmann M},
\textbf{Hollister EB}, \textbf{Lesniewski RA}, \textbf{Oakley BB},
\textbf{Parks DH}, \textbf{Robinson CJ}, \textbf{Sahl JW}, \textbf{Stres
B}, \textbf{Thallinger GG}, \textbf{Horn DJV}, \textbf{Weber CF}. 2009.
Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and Environmental Microbiology
\textbf{75}:7537--7541.
doi:\href{https://doi.org/10.1128/aem.01541-09}{10.1128/aem.01541-09}.}

\leavevmode\hypertarget{ref-Morris2013}{}%
\CSLLeftMargin{3. }
\CSLRightInline{\textbf{Morris A}, \textbf{Beck JM}, \textbf{Schloss
PD}, \textbf{Campbell TB}, \textbf{Crothers K}, \textbf{Curtis JL},
\textbf{Flores SC}, \textbf{Fontenot AP}, \textbf{Ghedin E},
\textbf{Huang L}, \textbf{Jablonski K}, \textbf{Kleerup E},
\textbf{Lynch SV}, \textbf{Sodergren E}, \textbf{Twigg H}, \textbf{Young
VB}, \textbf{Bassis CM}, \textbf{Venkataraman A}, \textbf{Schmidt TM},
\textbf{Weinstock GM}. 2013. Comparison of the respiratory microbiome in
healthy nonsmokers and smokers. American Journal of Respiratory and
Critical Care Medicine \textbf{187}:1067--1075.
doi:\href{https://doi.org/10.1164/rccm.201210-1913oc}{10.1164/rccm.201210-1913oc}.}

\leavevmode\hypertarget{ref-Bokulich2013}{}%
\CSLLeftMargin{4. }
\CSLRightInline{\textbf{Bokulich NA}, \textbf{Subramanian S},
\textbf{Faith JJ}, \textbf{Gevers D}, \textbf{Gordon JI}, \textbf{Knight
R}, \textbf{Mills DA}, \textbf{Caporaso JG}. 2013. Quality-filtering
vastly improves diversity estimates from illumina amplicon sequencing.
Nature Methods \textbf{10}:57--59.
doi:\href{https://doi.org/10.1038/nmeth.2276}{10.1038/nmeth.2276}.}

\leavevmode\hypertarget{ref-Fadrosh2014}{}%
\CSLLeftMargin{5. }
\CSLRightInline{\textbf{Fadrosh DW}, \textbf{Ma B}, \textbf{Gajer P},
\textbf{Sengamalay N}, \textbf{Ott S}, \textbf{Brotman RM},
\textbf{Ravel J}. 2014. An improved dual-indexing approach for
multiplexed 16S {rRNA} gene sequencing on the illumina {MiSeq} platform.
Microbiome \textbf{2}:6.
doi:\href{https://doi.org/10.1186/2049-2618-2-6}{10.1186/2049-2618-2-6}.}

\leavevmode\hypertarget{ref-Holm2019}{}%
\CSLLeftMargin{6. }
\CSLRightInline{\textbf{Holm JB}, \textbf{Humphrys MS}, \textbf{Robinson
CK}, \textbf{Settles ML}, \textbf{Ott S}, \textbf{Fu L}, \textbf{Yang
H}, \textbf{Gajer P}, \textbf{He X}, \textbf{McComb E}, \textbf{Gravitt
PE}, \textbf{Ghanem KG}, \textbf{Brotman RM}, \textbf{Ravel J}. 2019.
Ultrahigh-throughput multiplexing and sequencing of \(>\)500-base-pair
amplicon regions on the {Illumina} {HiSeq} 2500 platform. {mSystems}
\textbf{4}.
doi:\href{https://doi.org/10.1128/msystems.00029-19}{10.1128/msystems.00029-19}.}

\leavevmode\hypertarget{ref-Thompson2017}{}%
\CSLLeftMargin{7. }
\CSLRightInline{\textbf{Thompson LR}, \textbf{Jon G. Sanders and},
\textbf{McDonald D}, \textbf{Amir A}, \textbf{Ladau J}, \textbf{Locey
KJ}, \textbf{Prill RJ}, \textbf{Tripathi A}, \textbf{Gibbons SM},
\textbf{Ackermann G}, \textbf{Navas-Molina JA}, \textbf{Janssen S},
\textbf{Kopylova E}, \textbf{Vázquez-Baeza Y}, \textbf{González A},
\textbf{Morton JT}, \textbf{Mirarab S}, \textbf{Xu ZZ}, \textbf{Jiang
L}, \textbf{Haroon MF}, \textbf{Kanbar J}, \textbf{Zhu Q}, \textbf{Song
SJ}, \textbf{Kosciolek T}, \textbf{Bokulich NA}, \textbf{Lefler J},
\textbf{Brislawn CJ}, \textbf{Humphrey G}, \textbf{Owens SM},
\textbf{Hampton-Marcell J}, \textbf{Berg-Lyons D}, \textbf{McKenzie V},
\textbf{Fierer N}, \textbf{Fuhrman JA}, \textbf{Clauset A},
\textbf{Stevens RL}, \textbf{Shade A}, \textbf{Pollard KS},
\textbf{Goodwin KD}, \textbf{Jansson JK}, \textbf{Gilbert JA},
\textbf{Knight R}. 2017. A communal catalogue reveals earth's multiscale
microbial diversity. Nature \textbf{551}:457--463.
doi:\href{https://doi.org/10.1038/nature24621}{10.1038/nature24621}.}

\leavevmode\hypertarget{ref-McDonald2018}{}%
\CSLLeftMargin{8. }
\CSLRightInline{\textbf{McDonald D}, \textbf{Hyde E}, \textbf{Debelius
JW}, \textbf{Morton JT}, \textbf{Gonzalez A}, \textbf{Ackermann G},
\textbf{Aksenov AA}, \textbf{Behsaz B}, \textbf{Brennan C}, \textbf{Chen
Y}, \textbf{Goldasich LD}, \textbf{Dorrestein PC}, \textbf{Dunn RR},
\textbf{Fahimipour AK}, \textbf{Gaffney J}, \textbf{Gilbert JA},
\textbf{Gogul G}, \textbf{Green JL}, \textbf{Hugenholtz P},
\textbf{Humphrey G}, \textbf{Huttenhower C}, \textbf{Jackson MA},
\textbf{Janssen S}, \textbf{Jeste DV}, \textbf{Jiang L}, \textbf{Kelley
ST}, \textbf{Knights D}, \textbf{Kosciolek T}, \textbf{Ladau J},
\textbf{Leach J}, \textbf{Marotz C}, \textbf{Meleshko D}, \textbf{Melnik
AV}, \textbf{Metcalf JL}, \textbf{Mohimani H}, \textbf{Montassier E},
\textbf{Navas-Molina J}, \textbf{Nguyen TT}, \textbf{Peddada S},
\textbf{Pevzner P}, \textbf{Pollard KS}, \textbf{Rahnavard G},
\textbf{Robbins-Pianka A}, \textbf{Sangwan N}, \textbf{Shorenstein J},
\textbf{Smarr L}, \textbf{Song SJ}, \textbf{Spector T}, \textbf{Swafford
AD}, \textbf{Thackray VG}, \textbf{Thompson LR}, \textbf{Tripathi A},
\textbf{Vázquez-Baeza Y}, \textbf{Vrbanac A}, \textbf{Wischmeyer P},
\textbf{Wolfe E}, \textbf{Zhu Q}, \textbf{Knight R}, \textbf{Mann AE},
\textbf{Amir A}, \textbf{Frazier A}, \textbf{Martino C},
\textbf{Lebrilla C}, \textbf{Lozupone C}, \textbf{Lewis CM},
\textbf{Raison C}, \textbf{Zhang C}, \textbf{Lauber CL},
\textbf{Warinner C}, \textbf{Lowry CA}, \textbf{Callewaert C},
\textbf{Bloss C}, \textbf{Willner D}, \textbf{Galzerani DD},
\textbf{Gonzalez DJ}, \textbf{Mills DA}, \textbf{Chopra D},
\textbf{Gevers D}, \textbf{Berg-Lyons D}, \textbf{Sears DD},
\textbf{Wendel D}, \textbf{Lovelace E}, \textbf{Pierce E},
\textbf{TerAvest E}, \textbf{Bolyen E}, \textbf{Bushman FD}, \textbf{Wu
GD}, \textbf{Church GM}, \textbf{Saxe G}, \textbf{Holscher HD},
\textbf{Ugrina I}, \textbf{German JB}, \textbf{Caporaso JG},
\textbf{Wozniak JM}, \textbf{Kerr J}, \textbf{Ravel J}, \textbf{Lewis
JD}, \textbf{Suchodolski JS}, \textbf{Jansson JK},
\textbf{Hampton-Marcell JT}, \textbf{Bobe J}, \textbf{Raes J},
\textbf{Chase JH}, \textbf{Eisen JA}, \textbf{Monk J}, \textbf{Clemente
JC}, \textbf{Petrosino J}, \textbf{Goodrich J}, \textbf{Gauglitz J},
\textbf{Jacobs J}, \textbf{Zengler K}, \textbf{Swanson KS},
\textbf{Lewis K}, \textbf{Mayer K}, \textbf{Bittinger K}, \textbf{Dillon
L}, \textbf{Zaramela LS}, \textbf{Schriml LM}, \textbf{Dominguez-Bello
MG}, \textbf{Jankowska MM}, \textbf{Blaser M}, \textbf{Pirrung M},
\textbf{Minson M}, \textbf{Kurisu M}, \textbf{Ajami N}, \textbf{Gottel
NR}, \textbf{Chia N}, \textbf{Fierer N}, \textbf{White O}, \textbf{Cani
PD}, \textbf{Gajer P}, \textbf{Strandwitz P}, \textbf{Kashyap P},
\textbf{Dutton R}, \textbf{Park RS}, \textbf{Xavier RJ}, \textbf{Mills
RH}, \textbf{Krajmalnik-Brown R}, \textbf{Ley R}, \textbf{Owens SM},
\textbf{Klemmer S}, \textbf{Matamoros S}, \textbf{Mirarab S},
\textbf{Moorman S}, \textbf{Holmes S}, \textbf{Schwartz T},
\textbf{Eshoo-Anton TW}, \textbf{Vigers T}, \textbf{Pandey V},
\textbf{Treuren WV}, \textbf{Fang X}, \textbf{Xu ZZ}, \textbf{Jarmusch
A}, \textbf{Geier J}, \textbf{Reeve N}, \textbf{Silva R},
\textbf{Kopylova E}, \textbf{Nguyen D}, \textbf{Sanders K},
\textbf{Benitez RAS}, \textbf{Heale AC}, \textbf{Abramson M},
\textbf{Waldispühl J}, \textbf{Butyaev A}, \textbf{Drogaris C},
\textbf{Nazarova E}, \textbf{Ball M}, \textbf{Gunderson B}. 2018.
American gut: An open platform for citizen science microbiome research.
{mSystems} \textbf{3}.
doi:\href{https://doi.org/10.1128/msystems.00031-18}{10.1128/msystems.00031-18}.}

\leavevmode\hypertarget{ref-Schloss2011}{}%
\CSLLeftMargin{9. }
\CSLRightInline{\textbf{Schloss PD}, \textbf{Gevers D}, \textbf{Westcott
SL}. 2011. Reducing the effects of {PCR} amplification and sequencing
artifacts on 16S {rRNA}-based studies. {PLoS} {ONE} \textbf{6}:e27310.
doi:\href{https://doi.org/10.1371/journal.pone.0027310}{10.1371/journal.pone.0027310}.}

\leavevmode\hypertarget{ref-Schloss2018}{}%
\CSLLeftMargin{10. }
\CSLRightInline{\textbf{Schloss PD}. 2018. Identifying and overcoming
threats to reproducibility, replicability, robustness, and
generalizability in microbiome research. {mBio} \textbf{9}.
doi:\href{https://doi.org/10.1128/mbio.00525-18}{10.1128/mbio.00525-18}.}

\leavevmode\hypertarget{ref-Sogin2006}{}%
\CSLLeftMargin{11. }
\CSLRightInline{\textbf{Sogin ML}, \textbf{Morrison HG}, \textbf{Huber
JA}, \textbf{Welch DM}, \textbf{Huse SM}, \textbf{Neal PR},
\textbf{Arrieta JM}, \textbf{Herndl GJ}. 2006. Microbial diversity in
the deep sea and the underexplored "rare biosphere". Proceedings of the
National Academy of Sciences \textbf{103}:12115--12120.
doi:\href{https://doi.org/10.1073/pnas.0605127103}{10.1073/pnas.0605127103}.}

\leavevmode\hypertarget{ref-Huse2010}{}%
\CSLLeftMargin{12. }
\CSLRightInline{\textbf{Huse SM}, \textbf{Welch DM}, \textbf{Morrison
HG}, \textbf{Sogin ML}. 2010. Ironing out the wrinkles in the rare
biosphere through improved {OTU} clustering. Environmental Microbiology
\textbf{12}:1889--1898.
doi:\href{https://doi.org/10.1111/j.1462-2920.2010.02193.x}{10.1111/j.1462-2920.2010.02193.x}.}

\leavevmode\hypertarget{ref-Reeder2009}{}%
\CSLLeftMargin{13. }
\CSLRightInline{\textbf{Reeder J}, \textbf{Knight R}. 2009. The {{}}rare
biosphere{{}}: A reality check. Nature Methods \textbf{6}:636--637.
doi:\href{https://doi.org/10.1038/nmeth0909-636}{10.1038/nmeth0909-636}.}

\leavevmode\hypertarget{ref-Reid2011}{}%
\CSLLeftMargin{14. }
\CSLRightInline{\textbf{Reid A}, \textbf{Buckley M}. 2011. The rare
biosphere.}

\end{CSLReferences}

@@

\includegraphics{figure_1.png}

\textbf{Figure 1. Singletons are more common in samples with fewer
seqeunces and tend to be shared with samples having more sequences.} For
each of the 12 datasets, Spearman correlation coefficients were
calcualted between the number of sequences in each sample and the number
of singletons in the sample (A) and the fraction of its singletons that
were shared with another sample (C). Those correlations that were not
statisically significant had a P-value greater than 0.05. The faction of
singletons shared across samples (B) were calculated for each dataset.
The median value is shown with a solid circle and the 95\% confidence
interval is indicated by the solid line.

@@

\includegraphics{figure_2.png}

\textbf{Figure 2. Removing rare sequences from samples alters their
representation of alpha-diversity using amplicon sequence variants
(ASVs).} The average difference in the richness (A), Shannon diversity
(B), and Kullback-Leiber divergence (C) for each sample within a dataset
was calculated between the original community structures relative to
applying different minimum abundance thresholds.

\includegraphics{figure_3.png}

\textbf{Figure 3. Removing rare sequences from samples increases the
inter-sample variation for amplicon sequence variants (ASVs).} The
coefficient of variation in richness (A), Shannon diversity (B), and
Bray-Curtis distances (C) for each dataset was calculated using the null
distributed samples for each dataset with varying minimum abundance
thresholds.

\includegraphics{figure_4.png}

\textbf{Figure 4. Removing rare sequences from samples reduces the
statistical power to detect differences between empirically generated
treatment groups when using amplicon sequence variants (ASVs).} The
fraction of significant tests comparing the richness (A) and Shannon
diversity (B) using a Wilcox test and Bray-Curtis distances (C) using
analysis of molecular variance for each dataset was calculated using
empirically generated treatment groups containing equal numbers of
samples for each dataset with varying minimum abundance thresholds. For
each dataset and minimum abundance threshold, 100 randomizations were
peformed.

\includegraphics{figure_5.png}

\textbf{Figure 5. Removing rare sequences does not impact the false
detection rate unless the number of sequences per sample is confounded
with the treatment groups when using amplicon sequence variants (ASVs).}
The fraction of significant tests comparing the richness and Shannon
diversity using a Wilcox test and Bray-Curtis distances using analysis
of molecular variance for each dataset was calculated. Empirically
generated treatment groups were generated containing equal numbers of
samples where the samples represented a null distribution. In one
simulation the samples were randomly assigned to a treatment group (A)
and in the other the samples were assigned based on the number of
sequences in each sample (B). For each dataset and minimum abundance
threshold, 100 randomizations were peformed.

\includegraphics{figure_s1.png}

\textbf{Figure S1. Distribution of the number of sequences per sample in
the 12 datasets included in this study.} A different minimum number of
sequences per sample threshold was applied to each dataset based on
identifying natural breaks in the distribution of the number of
sequences per sample.

\includegraphics{figure_s2.png}

\textbf{Figure S2. Removing rare sequences from samples alters their
representation of alpha-diversity using operational taxonomic units
(OTUs).} The average difference in the richness (A), Shannon diversity
(B), and Kullback-Leiber divergence (C) for each sample within a dataset
was calculated between the original community structures relative to
applying different minimum abundance thresholds.

\includegraphics{figure_s3.png}

\textbf{Figure S3. Removing rare sequences from samples alters their
representation of alpha-diversity when regenerating samples using a null
distribution for each dataset.} The average difference in the richness,
Shannon diversity, and Kullback-Leiber divergence for each sample within
a dataset was calculated between the original community structures
relative to applying different minimum abundance thresholds. Some values
of Kullback-Leiber divergence are missing because undefined values were
calculated due to the removal of rare sequences. Data are shown for
amplicon sequence variants (ASVs) and operational taxonomic units (OTUs)

\includegraphics{figure_s4.png}
